A dispatch-based approach to fully allocate costs in optimal power systems

### 2020-08-28 paper/flow_based.tex
adjust numerical example section, but (!) change it now to only ap net

\title{Flow-Based Cost Allocation in Linear Optimized Power Systems are non-unique but localizable}
\author{F. Hofmann, Brown T.}

@@

\subsection*{Highlights}
\begin{itemize}
 \item The LMP always breaks down into contributions of different cost terms (OPEX, CAPEX, etc.) of the system.
 \item Breaking the LMP further down into contributions of components (generator, transmission line, etc.) requires assumptions on power flow assignments.
 \item Methods like Average Participation or Flow Based Market Coupling allow for localizing the P2P cost assignments and prioritize assignments within one market zone.
\end{itemize}

@@

In long-term operation and investment planning models, the total costs $\totalcost$ of a power system is the sum of mutliple cost terms. Typically, these include operational expenditures (OPEX) $\opexgeneration$ for generators, expenditures for emissions $\emissioncost$, capital expenditures (CAPEX) for generators $\capexgeneration$, CAPEX for the transmission system $\capexflow$ and possible other terms, \ie
\begin{align}
\totalcost &= \opexgeneration + \emissioncost +  \capexgeneration +  \capexflow + ...
\label{eq:total_cost}
\end{align}
In turn, each of these terms $\mathcal{C} = \{\opexgeneration, \emissioncost, \capexgeneration, \capexflow, ...\}$ consists of cost associated to the asset $i$ in the system, 
\begin{align}
    \cost = \sum_{i} \cost_{i}
    \label{eq:asset_cost}
\end{align} 
where an asset describes any component of the network, such as a generator, line, energy storage etc. %For example the total OPEX, $\opexgeneration = \sum_{s} \opexgeneration_{s}$, is the combined OPEX of all generators $s$.

In a cost-optimal setup with minimized $\totalcost$, the Locational Marginal Price (LMP) describes the price for an incremental increase of electricity demand $\demand$ at node $n$. It is given by the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$
\begin{align}
\lmp = \pdv{\,\totalcost}{\,\demand}
\label{eq:lmp}
\end{align}
% 
\begin{figure}[h]
\centering
\includegraphics[width=.8\linewidth]{price_decomposition.png}
\caption{Schematic decomposition of the Locational Market Price $\lmp$. In power system model with optimal long-term operation and planning, the total system costs $\totalcost$ split into different cost terms, \ie OPEX and CAPEX for production and transmission and possibly other expenditures. }
\label{fig:price_decomposition}
\end{figure}
% 
This leads to a nodal pricing where over the span of optimized timesteps $t$, the system costs are partially or totally payed back by the consumers 
\begin{align}
\totalcost - \remainingcost &=  \sum_{n,t} \lmp \, \demand
\label{eq:total_revenue}
\end{align}
depending on the costs $\remainingcost$ which are independent of the nodal demand  
\begin{align}
 \pdv{\remainingcost}{\demand} = 0
\end{align}
% 
Generally speaking, the cost term $\remainingcost$, not covered by the consumers, results from additional demands on the network design, such as capacity expansion limits or minimum share of one technology in the power mix. However, in most cases, where $\remainingcost \ll  \totalcost$, these play a minor role.

@@

% where $\costfactor$ is either a fixed operational price or a composition of shadowprices
Following the implications of \cref{eq:lmp,eq:total_revenue},  we define the cost $\allocatecost$ that consumers at bus $n$ have pay to asset $i$ at time $t$, in order to compensate for $\cost_i$, as  
\begin{align}
    \allocatecost = \costfactor \,\pdv{\state}{\demand} \demand
    \label{eq:cost_allocation}
\end{align}
The product on the right hand side 
\begin{align}
    \allocatestate = \pdv{\state}{\demand} \demand
    \label{eq:state_allocation}
\end{align}    
may be interpreted as the amount of power that asset $i$ supplies demand $\demand$ with. It heavily relies on the derivative of the operational state with respect to the nodal demand $\partial \state / \partial \demand$.
From the natural fact that the sum of all contributions must return the cost term, 
\begin{align}
    \cost_i = \sum_{n,t} \allocatecost
    \label{eq:cost_payback}
\end{align}
it follows that $\allocatestate$ must fulfill
\begin{align}
    \state = \sum_n \allocatestate
    \label{eq:state_allocation_constraint}
\end{align}
The last equation states that all power produced or processed by asset $i$ must be totally consumed by the network demand $\demand$.  
Finally, the total contribution from node $n$ at time $t$ to the cost term $\cost$ amounts 
\begin{align}
    \cost_{n,t} = \sum_i \allocatecost
\end{align}
\end{subequations}

@@

The transmission capacity $\capacityflow$ limits the flow $\flow$ in both directions,
\begin{align}
\flow - \capacityflow &\le 0 \resultsin{\muupperflow} \Forall{\ell,t} 
\label[constraint]{eq:upper_flow_capacity_constraint} \\
- \flow - \capacityflow &\le 0 \resultsin{\mulowerflow} \Forall{\ell,t} 
\label[constraint]{eq:lower_flow_capacity_constraint}
\end{align}
which yield the shadow prices $\muupperflow$ and $\mulowerflow$. Again, we use the result of \cite{brown_decreasing_2020} (for details see \cref{sec:zero_profit_flow}) which derives that over the whole time span, the investment in line $\ell$ is payed back by the shadow prices times the flow 
\begin{align}
\capexflow_\ell = \capitalpriceflow \capacityflow = \sum_{t} \left( \muupperflow - \mulowerflow \right)  \flow 
\label{eq:no_profit_capex_flow}
\end{align}
From here, we follow the scheme in \cref{eq:general_scheme} which finally defines the CAPEX allocation as 
\begin{align}
    \allocatecapexflow &=  
   \left( \muupperflow - \mulowerflow\right) \, \allocateflow
   \label{eq:allocate_capexFlow_detailed}
\end{align}
The quantity 
\begin{align}
 \allocateflow =  \pdv{\flow}{\demand}\, \demand
\end{align}
can be interpreted as the flow that the demand at node $n$ and time $t$ causes on line $\ell$.
% According to \cref{eq:state_allocation_constraint} it has to fulfill 
% \begin{align}
%  \flow = \sum_{n} \allocateflow
% \label{eq:allocate_flow_constraint}
% \end{align}
The  shadowprices $\muupperflow$ and $\mulowerflow$ again can be seen as a measure for necessity of transmission investments at $\ell$ at time $t$. Hence, the definition of $\allocatecapexflow$ states that consumers, which retrieve power flowing on congested lines, yielding a bound \cref{eq:upper_flow_capacity_constraint} or \eqref{eq:lower_flow_capacity_constraint}, pay compensations for the resulting investments at $\ell$. Again the sum of all CAPEX payments to line $\ell$ equals the total CAPEX spent. This is seen when summing \cref{eq:allocate_capexFlow_detailed} over all buses and time steps and using \cref{eq:state_allocation_constraint,eq:no_profit_capex_flow}

@@

In more realistic setups, generators, lines or other assets can only be built up to a certain limit. This might be due to land use restrictions or social acceptance problems. %In the following give a general solution how the cost allocation may handle this.  
However, when constraining the capacity $\capacity$  for a subset $I$ of assets to an upper limit $\capacityupper$, in the form of 
\begin{align}
    \capacity - \capacityupper \le 0 \resultsin{\muuppernom} \Forall{i \in I}
\label{eq:capacityexpansionmaximum},
\end{align}
the zero profit condition alters as soon as the constraint becomes binding. Then, the revenue of asset $i$ exceeds its total expenditures (OPEX + CAPEX). More precisely, the allocated CAPEX in \cref{tab:cost_allocation_map} will surpass the actual CAPEX of asset $i$ by the cost it has to pay for the scarcity, given by the absolute value of 
\begin{align}
    \scarcitycost_i = - \muuppernom \capacity \Forall{i \in I}
\end{align}
% \begin{align}
%  \cost_i - \remainingcost_i =  \sum_t \costfactor \, \state \Forall{i}
%  \label{eq:capex_generation_duality_bf2}
% \end{align}

@@

% Again, we can balance this effect out by updating the cost factor to 
% \begin{align}
%     \costfactor \rightarrow \costfactor +  \dfrac{\mulowernom \capacity}{\sum_t \state}
% \end{align}
% In constrast to \cref{eq:adjustment_lowernom} the price is not weighted by a factor but shifted for all time steps. This is required as 
% As $\mulowernom$ is a positive number, this will increase the price and therefore the revenue for asset $i$.

@@

\item Average Participation (AP) \cite{bialek_tracing_1996,achayuthakan_electricity_2010-1} which traces the flow from producer to consumer following the law of proportional sharing. \label{ap}
\item Flow Based Market Coupling (FBMC) which uses zonal PTDF for allocating power within predefined regions. The interregional exchange is only allocating net power decifit or excess of the regions. \label{fbmc}

@@

Further, as literature has often pointed out, the EBE principal \ref{ebe} does not suit for large networks where remote buses would interconnect in the same way as buses in close vicinity \cite{gil_multiarea_2005}. The AP based type \ref{ap} tackles this problem by restricting P2P trades to those which are traceable when applying the proportional sharing principal. Therefore $\allocategeneration$ denotes that part of power produced by bus $m$ which, when only following in the direction of $\flow$, ends up at bus $n$. Type \ref{fbmc} further allows to control the regions or market zones which are netted out in a first step. If in a region $R$ the generation undercuts the demand, $\sum_{n \in R} \nodalgeneration \le \sum_{n \in R} \demand$, none of the inner-regional generation is assigned to other regions.

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=\linewidth]{example_network.png}
    \caption{Illustrative example of a 2-bus network with one optimized time step. Fixed prices and constraining values are given in the left box for each bus and the transmission line. Optimized values are given in the right boxes. Bus~1 has a cheaper operational price $o$, capital prices are the same for both. As both generator capacities are constraint to 100~MW, the optimization also deploys the generator at bus~2. The resulting electricity prices $\lambda$ are then a composition of all prices for operation and capital investments.}
    \label{fig:example_network}
    \end{figure*}
In the following, we illustrate the cost allocation and explain its functionality by taking the example of a two bus system with one optimized time step, shown in \cref{fig:example_network}. \\
The two buses are connected via one transmission line, each has one generator. Whereas generator~1 (at bus~1) has an operational price of 50 \euro/\megawatthour, generator~2 (at bus~2) has a higher operational price of 200~\euro/\megawatthour. For both the CAPEX rate is set to 500~\euro/MW and the maxmimal capacity is limited to $\capacitygenerationupper$~=~100~MW. The transmission line has a CAPEX rate of 100~\euro/MW and no upper capacity limit. With a demand of 60~MW at bus~1 and 90~MW at bus~2, the optimization expands the cheaper generator at bus~1 to its full limit of 100~MW. The 40~MW excess power, not consumed at bus~1, flows to bus~2 where the generator is built with only 50~MW.  
% \begin{figure}[h]
%     \includegraphics[width=\linewidth]{example_allocation_bus1_gross_ebe.png}
%     \label{fig:example_allocation_bus1}
%     \caption{Power allocations of equivalently allocated gross power, type \ref{gross}\ref{ebe}, for bus~1 (a) and bus~2 (b) of the example network in \cref{fig:example_network}. Bus~1 retrieves 40~MW from itself and 20~MW from bus~2.}
% \end{figure}
% % 
% \begin{figure}[b]
%     \includegraphics[width=\linewidth]{example_allocation_bus2_gross_ebe.png}
%     \label{fig:example_allocation_bus2_gross_ebe}
%     \caption{Type \ref{gross}\ref{ebe} power allocation for bus~2 of the example network in \cref{fig:example_network}. Bus~2 retrieves 60~MW from bus~1 and self supplies 30~MW. The sum of both net flows equals the resulting flow of $f_1=40$~MW.}
% \end{figure}
\begin{figure*}[h!]
    \begin{subfigure}[c]{.495\linewidth}
    \includegraphics[width=\linewidth]{example_allocation_bus1_gross_ebe.png}
    \vspace{-40pt}
    \subcaption{}
    \label{fig:example_allocation_bus1}
    \end{subfigure}
    \begin{subfigure}[c]{.495\linewidth}
    \includegraphics[width=\linewidth]{example_allocation_bus2_gross_ebe.png}
    \vspace{-40pt}
    \subcaption{}
    \label{fig:example_allocation_bus2_gross_ebe}
    \end{subfigure}
    \caption{Power allocations of equivalently allocated gross power, type \ref{gross}\ref{ebe}, for bus~1 (a) and bus~2 (b) of the example network in \cref{fig:example_network}. Bus~1 retrieves 40~MW from itself and 20~MW from bus~2. The latter in turn retrieves 60~MW from bus~1 and self supplies 30~MW.
    The sum of both net flows equals the resulting flow of $f_1=40$~MW. \vspace{-0pt}}
    \label{fig:example_allocation}
\end{figure*}
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{example_payoff_gross_ebe.png}
    \caption{Full P2P cost allocation for the example setup shown in \cref{fig:example_network}. The payments are derived on the basis of \cref{eq:allocate_opexGeneration_detailed,eq:allocate_capexGeneration_detailed,eq:allocate_capexFlow_detailed}. Consumers at bus $n$ have to pay each generator proportional to their consumption. As we only consider one time step the proportionality applies for OPEX $\allocateopex$ and CAPEX $\mathcal{C}_{n \rightarrow i}$. As bus~1 induces a relieving flow an line~1 and therefore ``prevents`` further tranmission expansion, it is rewarded proportional to the relief.}
    \label{fig:example_payoff}
\end{figure}    
\subsubsection*{Allocating Gross Power using EBE (\ref{gross}\ref{ebe}) }

\Cref{fig:example_allocation} shows the allocated transactions on basis of equvalently traded gross power (type \ref{gross}\ref{ebe}) for both buses 1~\&~2 separately. The resulting P2P payments are given in \cref{fig:example_payoff}.
The upper graph \cref{fig:example_allocation_bus1} shows that $A_{1 \rightarrow 1}=40$~MW at bus~1 are self-sustained. With only one generator at bus 1, consumers at bus~1 consequently pay 2k~\euro~OPEX and 22k~\euro~CAPEX to the generator~1. The remaining 20~MW come from bus~2 and induce a subflow on line 1 of $A_{\ell = 1, 1} = -20$. As this flow is in contrary direction to the total flow, it is relieving the transmission system. This translates to a congestion reward for consumers at bus~1 of $c_{\ell=1} A_{\ell = 1, 1}$~=~2k~\euro\, which is exactly the cost that had to be spent on the transmission system if bus~1 didn't induce a relieving flow, see again \cref{fig:example_payoff}.

The lower graph \cref{fig:example_allocation_bus2_gross_ebe} illustrates the impact of consumption at bus~2. As $d_2$ is higher than $d_1$ the retrievements from both generators are proportionately increased as well as the OPEX and CAPEX allocations to the generators. But instead of a relieving flow, consumers at bus~2 drive the burdening flow in direction of congestion. Hence the payoff to the transmission system is positive and much higher than for bus~1.

The sum of all rows in the payoff matrix in \cref{fig:example_payoff} yields the revenues of the assets $m, \ell$. These values match their overall spendings, \textit{e.g.} the total revenue of the transmission line is 4k~\euro\, which equals the cost for investments $c_{1}\,F_{1}$. The sum of all columns yields the total payment of consumers at bus $n$. For example the sum of payments of consumers at bus~1 is 36k~\euro. This is exactly the electricity price of 600~\euro/MW times the consumption of 60~MW, $\lambda_1 d_1$. \\

The fact that OPEX and CAPEX allocations are proportional to the total consumption at a bus results from optimizing one time step only. In larger optimization problems with multiple time steps the CAPEX allocation takes effect only for time steps in which one or more of the capacity contraints  \cref{eq:upper_generation_capacity_constraint,eq:lower_flow_capacity_constraint,eq:upper_flow_capacity_constraint} become binding.

\subsubsection*{Allocating Net Power using EBE (\ref{net}\ref{ebe}) }

\begin{figure}[h!]
    \centering
    \includegraphics[width=\linewidth]{example_payoff_net_ebe.png}
    \caption{Full P2P cost allocation for the example setup shown in \cref{fig:example_network} when equivalently allocation net power injection (scheme \ref{ebe}\ref{net}).}
    \label{fig:example_payoff_net_ebe}
    \end{figure}    
% 
In contrast the to equivalent allocation of gross power production, netting out injections for each bus leads to less P2P payments. The resulting payment given in \cref{fig:example_payoff_net_ebe} builds on the allocated power flow shown \cref{fig:example_allocation_net_ebe} in \cref{sec:example_plots}. As bus~2 does not produce excess power, none of its power production is assigned to bus~1 and thus no payment of bus~1 to bus~2 allocated. Neither has bus~1 to pay fee to the transmission system as it only exports power. So all its consumers pay to its local generators. Bus~2 in constrast bear all CAPEX for the transmission system as well as CAPEX and OPEX for generators at bus~1.
% 
Again the cummulative payments per bus meet the nodal spendings $\lmp \, \demand$. The cummulative revenues per generator and transmission line meet the all CAPEX and OPEX.

As the other schemes \ref{ap}\ref{gross} - \ref{fbmc}\ref{net} apart from the EBE scheme don't distinguishably modify allocations within the presented example, we move on to a more realistic setup where systemic and local effects show up.

@@

\subsection{CO$_2$ constraint}

@@

% \subsection{Brownfield Optimization and Capacity Restricitions}

@@

The allocation of gross generation to demands $\demand$ is straigthforwardly obained by a proportional distribution of the generation, \ie

>>>>

\title{A dispatch-based approach to fully allocate costs in optimal power systems}
\author{F. Hofmann, Brown T.}

@@

\subsection*{Highlights}
\begin{itemize}
 \item The LMP always breaks down into contributions of different cost terms (OPEX, CAPEX, etc.) of the system.
 \item Breaking the LMP further down into contributions of components (generator, transmission line, etc.) requires assumptions on power flow assignments, where prominent methods as Average Participation or Flow Based Market Coupling can be applied.
 \item ...
 \item ...
\end{itemize}

@@

In long-term operation and investment planning models, the total costs $\totalcost$ of a power system is the sum of multiple cost terms. Typically, these include operational expenditures (OPEX) $\opexgeneration$ for generators, expenditures for emissions $\emissioncost$, capital expenditures (CAPEX) for generators $\capexgeneration$, CAPEX for the transmission system $\capexflow$ and possible other terms, \ie
\begin{align}
\totalcost &= \opexgeneration + \emissioncost +  \capexgeneration +  \capexflow + ...
\label{eq:total_cost}
\end{align}
In turn, each of these terms $\mathcal{C} = \{\opexgeneration, \emissioncost, \capexgeneration, \capexflow, ...\}$ consists of cost associated to the asset $i$ in the system, 
\begin{align}
    \cost = \sum_{i} \cost_{i}
    \label{eq:asset_cost}
\end{align} 
where an asset describes any component of the network, such as a generator, line, energy storage etc. %For example the total OPEX, $\opexgeneration = \sum_{s} \opexgeneration_{s}$, is the combined OPEX of all generators $s$.

In a cost-optimal setup with minimized $\totalcost$, the Locational Marginal Price (LMP) describes the price for an incremental increase of electricity demand $\demand$ at node $n$. It is given by the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$
\begin{align}
\lmp = \pdv{\,\totalcost}{\,\demand}
\label{eq:lmp}
\end{align}
% 
\begin{figure}[h]
\centering
\includegraphics[width=.8\linewidth]{price_decomposition.png}
\caption{Schematic decomposition of the Locational Market Price $\lmp$. In power system model with optimal long-term operation and planning, the total system costs $\totalcost$ split into different cost terms, \ie OPEX and CAPEX for production and transmission and possibly other expenditures. }
\label{fig:price_decomposition}
\end{figure}
% 
This leads to a nodal pricing where over the span of optimized time steps $t$, the system costs are partially or totally payed back by the consumers 
\begin{align}
\totalcost - \remainingcost &=  \sum_{n,t} \lmp \, \demand
\label{eq:total_revenue}
\end{align}
depending on the costs $\remainingcost$ which are independent of the nodal demand  
\begin{align}
 \pdv{\remainingcost}{\demand} = 0
\end{align}
% 
Generally speaking, the cost term $\remainingcost$, not covered by the consumers, results from additional demands on the network design, such as capacity expansion limits or minimum share of one technology in the power mix. However, in most cases, where $\remainingcost \ll  \totalcost$, these play a minor role.

@@

% where $\costfactor$ is either a fixed operational price or a composition of shadow prices
Following the implications of \cref{eq:lmp,eq:total_revenue},  we define the cost $\allocatecost$ that consumers at bus $n$ have pay to asset $i$ at time $t$, in order to compensate for $\cost_i$. This leads us to 
\begin{align}
    \allocatecost = \costfactor \,\pdv{\state}{\demand} \demand
    \label{eq:cost_allocation}
\end{align}
The derivative on the right hand side is defined through the sensitivity of the operational variable $\state$ at the optimum against changes in the demand. 
\begin{align}
    \allocatestate \rightarrow\pdv{\state}{\demand} \demand
    \label{eq:state_allocation}
\end{align}    
... may be interpreted as the amount of power that asset $i$ supplies demand $\demand$ with. It heavily relies on the derivative of the operational state with respect to the nodal demand $\partial \state / \partial \demand$. 
From the natural fact that the sum of all contributions must return the cost term, 
\begin{align}
    \cost_i = \sum_{n,t} \allocatecost
    \label{eq:cost_payback}
\end{align}
it follows that $\allocatestate$ must fulfill
\begin{align}
    \state = \sum_n \allocatestate
    \label{eq:state_allocation_constraint}
\end{align}
The last equation states that all power produced or processed by asset $i$ must be totally consumed by the network demand $\demand$.  
Finally, the total contribution from node $n$ at time $t$ to the cost term $\cost$ amounts 
\begin{align}
    \cost_{n,t} = \sum_i \allocatecost
\end{align}
\end{subequations}

@@

The transmission capacity $\capacityflow$ limits the flow $\flow$ in both directions,
\begin{align}
\flow - \capacityflow &\le 0 \resultsin{\muupperflow} \Forall{\ell,t} 
\label[constraint]{eq:upper_flow_capacity_constraint} \\
- \flow - \capacityflow &\le 0 \resultsin{\mulowerflow} \Forall{\ell,t} 
\label[constraint]{eq:lower_flow_capacity_constraint}
\end{align}
which yield the shadow prices $\muupperflow$ and $\mulowerflow$. Again, we use the result of \cite{brown_decreasing_2020} (for details see \cref{sec:zero_profit_flow}) which derives that over the whole time span, the investment in line $\ell$ is payed back by the shadow prices times the flow 
\begin{align}
\capexflow_\ell = \capitalpriceflow \capacityflow = \sum_{t} \left( \muupperflow - \mulowerflow \right)  \flow 
\label{eq:no_profit_capex_flow}
\end{align}
From here, we follow the scheme in \cref{eq:general_scheme} which finally defines the CAPEX allocation as 
\begin{align}
    \allocatecapexflow &=  
   \left( \muupperflow - \mulowerflow\right) \, \allocateflow
   \label{eq:allocate_capexFlow_detailed}
\end{align}
The quantity 
\begin{align}
 \allocateflow =  \pdv{\flow}{\demand}\, \demand
\end{align}
can be interpreted as the flow that the demand at node $n$ and time $t$ causes on line $\ell$.
% According to \cref{eq:state_allocation_constraint} it has to fulfill 
% \begin{align}
%  \flow = \sum_{n} \allocateflow
% \label{eq:allocate_flow_constraint}
% \end{align}
The  shadow prices $\muupperflow$ and $\mulowerflow$ again can be seen as a measure for necessity of transmission investments at $\ell$ at time $t$. Hence, the definition of $\allocatecapexflow$ states that consumers, which retrieve power flowing on congested lines, yielding a bound \cref{eq:upper_flow_capacity_constraint} or \eqref{eq:lower_flow_capacity_constraint}, pay compensations for the resulting investments at $\ell$. Again the sum of all CAPEX payments to line $\ell$ equals the total CAPEX spent. This is seen when summing \cref{eq:allocate_capexFlow_detailed} over all buses and time steps and using \cref{eq:state_allocation_constraint,eq:no_profit_capex_flow}

@@

In more realistic setups, generators, lines or other assets can only be built up to a certain limit. This might be due to land use restrictions or social acceptance problems. %In the following give a general solution how the cost allocation may handle this.  
However, when constraining the capacity $\capacity$  for a subset $I$ of assets to an upper limit $\capacityupper$, in the form of 
\begin{align}
    \capacity - \capacityupper \le 0 \resultsin{\muuppernom} \Forall{i \in I}
\label{eq:capacityexpansionmaximum},
\end{align}
the zero profit condition alters as soon as the constraint becomes binding. Then, the revenue of asset $i$ exceeds its total expenditures (OPEX + CAPEX). More precisely, the allocated CAPEX in \cref{tab:cost_allocation_map} will surpass the actual CAPEX of asset $i$ by the cost it has to pay for the scarcity, given by the absolute value of 
\begin{align}
    \scarcitycost_i = - \muuppernom \capacity \Forall{i \in I}
    \label{eq:scarcitycost}
\end{align}
% \begin{align}
%  \cost_i - \remainingcost_i =  \sum_t \costfactor \, \state \Forall{i}
%  \label{eq:capex_generation_duality_bf2}
% \end{align}

@@

% Again, we can balance this effect out by updating the cost factor to 
% \begin{align}
%     \costfactor \rightarrow \costfactor +  \dfrac{\mulowernom \capacity}{\sum_t \state}
% \end{align}
% In contrast to \cref{eq:adjustment_lowernom} the price is not weighted by a factor but shifted for all time steps. This is required as 
% As $\mulowernom$ is a positive number, this will increase the price and therefore the revenue for asset $i$.

@@

\item Average Participation (AP) \cite{bialek_tracing_1996,achayuthakan_electricity_2010} which traces the flow from producer to consumer following the law of proportional sharing. \label{ap}
\item Flow Based Market Coupling (FBMC) which uses zonal PTDF for allocating power within predefined regions. The interregional exchange is only allocating net power deficit or excess of the regions. \label{fbmc}

@@

Further, as literature has often pointed out, the EBE principal \ref{ebe} does not suit for large networks where remote buses would interconnect in the same way as buses in close vicinity \cite{gil_multiarea_2005}. The AP based type \ref{ap} tackles this problem by restricting P2P trades to those which are traceable when applying the proportional sharing principal. Therefore $\allocategeneration$ denotes that part of power produced by bus $m$ which, when only following in the direction of $\flow$, ends up at bus $n$. Type \ref{fbmc} further allows to control the regions or market zones which are netted out in a first step. If in a region $R$ the generation undercuts the demand, $\sum_{n \in R} \nodalgeneration \le \sum_{n \in R} \demand$, none of the inner-regional generation is assigned to other regions. However, it relies on further assumptions such as the Generation Shift Keys determining the production which are deployed for the inter-regional exchange. 
From this point of view, we set the main focus of this research to the AP based scheme, as locality and little de-pendency on external decisions are strong arguments for a transparent cost allocation. However, we will use the EBE based cost allocation for the following example, in order to depict the functionality of the cost allocation and the fundamental difference between the allocation of gross and net power.

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=\linewidth]{example_network.png}
    \caption{Illustrative example of a 2-bus network with one optimized time step. Fixed prices and constraining values are given in the left box for each bus and the transmission line. Optimized values are given in the right boxes. Bus~1 has a cheaper operational price $o$, capital prices are the same for both. As both generator capacities are constraint to 100~MW, the optimization also deploys the generator at bus~2. The resulting electricity prices $\lambda$ are then a composition of all prices for operation and capital investments.}
    \label{fig:example_network}
    \end{figure*}
    % 
% 
% 
Consider a two bus system, as shown in \cref{fig:example_network}, with one transmission line and one generator per bus. Whereas generator~1 (at bus~1) has an operational price of 50 \euro/\megawatthour, generator~2 (at bus~2) has a higher operational price of 200~\euro/\megawatthour. For both the CAPEX rate is set to 500~\euro/MW and the maximal capacity is limited to $\capacitygenerationupper$~=~100~MW. The transmission line has a CAPEX rate of 100~\euro/MW and no upper capacity limit. With a demand of 60~MW at bus~1 and 90~MW at bus~2, the optimization expands the cheaper generator at bus~1 to its full limit of 100~MW. The 40~MW excess power, not consumed at bus~1, flows to bus~2 where the generator is built with only 50~MW.  
% 
% 
% \begin{figure}[h]
%     \includegraphics[width=\linewidth]{example_allocation_bus1_gross_ebe.png}
%     \label{fig:example_allocation_bus1}
%     \caption{Power allocations of equivalently allocated gross power, type \ref{gross}\ref{ebe}, for bus~1 (a) and bus~2 (b) of the example network in \cref{fig:example_network}. Bus~1 retrieves 40~MW from itself and 20~MW from bus~2.}
% \end{figure}
% % 
% \begin{figure}[b]
%     \includegraphics[width=\linewidth]{example_allocation_bus2_gross_ebe.png}
%     \label{fig:example_allocation_bus2_gross_ebe}
%     \caption{Type \ref{gross}\ref{ebe} power allocation for bus~2 of the example network in \cref{fig:example_network}. Bus~2 retrieves 60~MW from bus~1 and self supplies 30~MW. The sum of both net flows equals the resulting flow of $f_1=40$~MW.}
% \end{figure}
% 
\begin{figure*}[h!]
    \begin{subfigure}[c]{.495\linewidth}
    \includegraphics[width=\linewidth]{example_allocation_bus1_gross_ebe.png}
    \vspace{-40pt}
    \subcaption{}
    \label{fig:example_allocation_bus1}
    \end{subfigure}
    \begin{subfigure}[c]{.495\linewidth}
    \includegraphics[width=\linewidth]{example_allocation_bus2_gross_ebe.png}
    \vspace{-40pt}
    \subcaption{}
    \label{fig:example_allocation_bus2_gross_ebe}
    \end{subfigure}
    \caption{Power allocations for gross power injection using the EBE scheme, type \ref{gross}\ref{ebe}, for bus~1 (a) and bus~2 (b) of the example network in \cref{fig:example_network}. Bus~1 retrieves 40~MW from itself and 20~MW from bus~2. The latter in turn retrieves 60~MW from bus~1 and self supplies 30~MW.
    The sum of both net flows equals the resulting flow of $f_1=40$~MW.}
    \label{fig:example_allocation}
\end{figure*}
% 
% 
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{example_payoff_gross_ebe.png}
    \caption{Full P2P cost allocation for the example setup shown in \cref{fig:example_network}. The payments are derived on the basis of \cref{eq:allocate_opexGeneration_detailed,eq:allocate_capexGeneration_detailed,eq:allocate_capexFlow_detailed}. Consumers at bus $n$ have to pay each generator proportional to their consumption. As we only consider one time step the proportionality applies for OPEX $\allocateopex$ and CAPEX $\mathcal{C}_{n \rightarrow i}$. As bus~1 induces a relieving flow an line~1 and therefore ``prevents`` further transmission expansion, it is rewarded proportional to the relief.}
    \label{fig:example_payoff}
\end{figure}    
% 
% 
% 
\subsubsection*{Allocating Gross Power Production}

\Cref{fig:example_allocation} shows the allocated transactions on basis of gross power injection for both buses 1~\&~2 separately. The resulting P2P payments are given in \cref{fig:example_payoff}.
The upper graph \cref{fig:example_allocation_bus1} shows that $A_{1 \rightarrow 1}=40$~MW at bus~1 are self-sustained. With only one generator at bus 1, consumers at bus~1 consequently pay 2k~\euro~OPEX and 22k~\euro~CAPEX to the generator~1. The remaining 20~MW come from bus~2 and induce a subflow on line 1 of $A_{\ell = 1, 1} = -20$. As this flow is in contrary direction to the total flow, it is relieving the transmission system. This translates to a congestion reward for consumers at bus~1 of $c_{\ell=1} A_{\ell = 1, 1}$~=~2k~\euro\, which is exactly the cost that had to be spent on the transmission system if bus~1 didn't induce a relieving flow, see again \cref{fig:example_payoff}.

The lower graph \cref{fig:example_allocation_bus2_gross_ebe} illustrates the impact of consumption at bus~2. As $d_2$ is higher than $d_1$, the reception from both generators are proportionately increased as well as the OPEX and CAPEX allocations to the generators. But instead of a relieving flow, consumers at bus~2 drive the burdening flow in direction of congestion. Hence the payoff to the transmission system is positive and much higher than for bus~1.

The sum of all rows in the payoff matrix in \cref{fig:example_payoff} yields the revenues of the assets $m, \ell$. These values match their overall spending, \textit{e.g.} the total revenue of the transmission line is 4k~\euro\, which equals the cost for investments $c_{1}\,F_{1}$. The sum of all columns yields the total payment of consumers at bus $n$. For example the sum of payments of consumers at bus~1 is 36k~\euro. This is exactly the electricity price of 600~\euro/MW times the consumption of 60~MW, $\lambda_1 d_1$. \\

The fact that OPEX and CAPEX allocations are proportional to the total consumption at a bus results from optimizing one time step only. In larger optimization problems with multiple time steps the CAPEX allocation takes effect only for time steps in which one or more of the capacity constraints  \cref{eq:upper_generation_capacity_constraint,eq:lower_flow_capacity_constraint,eq:upper_flow_capacity_constraint} become binding.

\subsubsection*{Allocating Net Power Production}

\begin{figure}[h!]
    \centering
    \includegraphics[width=\linewidth]{example_payoff_net_ebe.png}
    \caption{Full P2P cost allocation for the example setup shown in \cref{fig:example_network} when allocating net power injection using the EBE scheme (type \ref{net}\ref{ebe}). This leads to less and more intuitive payments.}
    \label{fig:example_payoff_net_ebe}
    \end{figure}    
% 
In contrast the to equivalent allocation of gross power production, netting out injections for each bus leads to less P2P payments. The resulting payment given in \cref{fig:example_payoff_net_ebe} builds on the allocated power flow shown \cref{fig:example_allocation_net_ebe} in \cref{sec:example_plots}. As bus~2 does not produce excess power, none of its power production is assigned to bus~1 and thus no payment of bus~1 to bus~2 allocated. Neither has bus~1 to pay fee to the transmission system as it only exports power. So, consumers at bus~1 pay to its local generator. Bus~2 in contrast bear all CAPEX for the transmission system as well as CAPEX and OPEX for generators at bus~1. 
% 
Again the cumulative payments per bus meet the nodal spending $\lmp \, \demand$. The cumulative revenues per generator and transmission line meet the all CAPEX and OPEX. Note this gives the same result as when allocating net injection with the Average Participation \ref{net}\ref{ap}.

The example shows that allocating net power injection only, reduces the number of peer-to-peer payments significantly, which leads to a much clearer pictures. The same counts for the AP scheme which we will use in the following application case.

%We emphasize that the AP based allocation does not imply traceability of the resulting flows, as only the \textit{source}-\textit{sink} relations $\allocatepeer$ are determined by it, the flow allocation still follow the power plow equation given as shown in \cref{eq:allocate_peer_to_allocate_flow}.

@@

\subsection{Emission Constraint}

@@

% \subsection{Brownfield Optimization and Capacity Restrictions}

@@

The allocation of gross generation to demands $\demand$ is straightforwardly obtained by a proportional distribution of the generation, \ie
